Foregrounds in Wide--Field Redshifted 21~cm Power Spectra

### 2015-02-11 ms.tex
Address referee comments for new draft
from d946469:
============

In ApJ_referee_report_1_addressed.txt,

(*) Address referee comments with reference to the alternate version of
the draft
(*) Draft contents of the cover letter

In ms.tex,

(*) Change wording in abstract slightly
(*) Change wording in S5 slightly and cite Lonsdale et al. 2009 and
Tingay et al. 2013 for the MWA instrument.
(*) Change wording on Tsys estimation slightly in S5.2 Modeling
subsection
(*) Change wording slightly and mention Pober et al (2015, in
preparation) Foreground model subsection to support use of wide-field
f

Detection of 21~cm emission of H{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission. We investigate the signatures of wide--field measurements and an all--sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs. We demonstrate interferometric measurements are inherently sensitive to all scales, including the largest angular scales, owing to the nature of wide--field measurements. These wide--field effects are generic to all observations but antenna shapes impact their amplitudes substantially. A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array. Comparing data from recent Murchison Widefield Array observations, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. We identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings. For signals entering through the primary field of view, compact emission dominates the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier delay space. Based on these results, we propose that selective down--weighting of data based on antenna spacing and time can mitigate foreground contamination substantially by a factor $\sim 100$ with negligible loss of sensitivity.

@@

{\bf In \S\ref{sec:delay-spectrum} we provide a brief overview of the delay spectrum technique \citep{par12a,par12b}. We investigate signatures generic to all wide--field measurements of EoR power spectra in \S\ref{sec:wide-field}. In \S\ref{sec:sim}, we present the foreground model and a variety of instrument models to rank antenna shapes based on foreground contamination. In \S\ref{sec:MWA}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are compared with the data and analyzed for foreground signatures. We report two important findings: foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique that minimizes foreground contamination using prior knowledge of the sky to down--weight adversely contaminated baselines}. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

Interferometer array data known as {\it visibilities}, $V_b(f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis. If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_\textrm{i}(f)$ denotes instrumental bandpass weights, then $V_b(f)$ can be written as:
\begin{align}\label{eqn:obsvis}
  V_b(f) &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
\end{align}
where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector. This equation is valid {\bf in general, including wide--field measurements and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}}.

@@

In order to express a quantity derived from $\tilde{V}_b(\tau)$ that has units K$^2$(Mpc/$h$)$^3$ which is identical to that of the cosmological H{\sc i} power spectrum, we define the delay power spectrum:
\begin{align}\label{eqn:delay-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \\
  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
\end{align}
where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology.

@@

{\bf In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s. Thus with reference to the MWA, the {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade}.

@@

In wide--field measurements, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view. We use a foreground model that includes both diffuse and bright compact components.

@@

These criteria yield 100 objects from the SUMSS catalog and 250 objects from the NVSS catalog. Together with the diffuse foreground model, we obtain an all--sky foreground model consisting of both compact and diffuse emission. Figure~\ref{fig:sky-model-generic} shows the diffuse (top) and compact (bottom) foreground emission model used in our study. In this snapshot at 0.09~hours LST, the Galactic center in the diffuse model has just set in the West.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figure4.eps}
\includegraphics[width=\linewidth]{figure3.eps}
\caption{Foreground model at 185~MHz consisting of diffuse emission (top) and bright point sources (bottom) visible during a snapshot at 0.09~hours LST. In the diffuse model, the Galactic center appears to have just set in the West. \label{fig:sky-model-generic}}
\end{figure}

@@

The power patterns of the aforementioned antenna geometries at 185~MHz for the {\it zenith} pointing are shown in Figure~\ref{fig:power-patterns}.

@@

We now use our simulations to analyze features in observed delay power spectrum obtained using the MWA.

@@

MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013. The MWA is a 128--tile interferometer capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie. This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes. The array layout is described in \citet{bea12}. %{\bf At the observing frequency of 185~MHz, we use $A_\textrm{e}=10.5$~m$^2$}.

@@

{\bf The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{bow06,bea13,thy13}. Over 600~hours have been currently observed using the MWA, targeting the science objectives outlined in \citet{bow13}.}

The MWA targets two primary low-foreground fields for reionization observations. Here, we focus on the field at RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$. The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of the dipoles in an MWA tile. The pointing system is optimized to points on a regular $\sim$7\arcdeg~ grid. The sky drifts across to the nearest available pointing, shifting between grid points (once every $\sim 30$~minutes). This process is repeated throughout the course of the observation $\approx 4.86$~hours.

The observations {\bf used} here were acquired on 2013~August~23. We have chosen two, 112 second long sections from this night for detailed study. These were chosen to provide a selection of possible foreground and instrumental conditions. As an example of a nominal observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hours from zenith. This pointing has a significantly higher secondary lobe structure and is observed when the bright galactic center is well above the horizon.

@@

{\bf We estimate thermal noise, $\Delta V$, in the visibilities, $V_b(f)$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
\begin{align}\label{eqn:Tsys}
  \Delta \tilde{V} &= \sqrt{N_\textrm{ch}}\,\Delta V \Delta f, \quad\textrm{and} \\
  \Delta V &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}}{A_\textrm{e}\sqrt{2\,\Delta f\,\Delta t}},
\end{align}
where, $\Delta f = 80$~kHz, $\Delta t = 112$~s, and $N_\textrm{ch}=\Delta B/\Delta f$ is the number of frequency channels. The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus yields a robust estimate of $T_\textrm{sys}$. We find the average system temperature to be $\sim 95$~K. Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.}

@@

Having shown that the simulation matches the data to the level of expected uncertainties, we proceed to examine in further detail the key signatures seen in simulated delay spectra. A number of factors are responsible for the characteristics noted in the delay spectra obtained from data and through simulations. In subsequent sections, we provide a detailed explanation of our results as a combination of these factors. Note that numerous features may overlap at different degrees of significance depending on combinations of parameters. We assign the features to their predominant causes. Secondly, we have used noiseless cases to clearly illustrate the observed foreground signatures. With the addition of noise in the visibilities, some of the weaker features may not be as prominently visible. Since the foreground signatures are far too numerous {\bf and subject to a multitude of parameters like} baseline length and orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features in the delay spectra of foreground emission.

@@

The most prominent signature seen in the {\it off--zenith} pointing (top left panel of Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern. It appears as a bright branch near the negative delay horizon delay limit. This feature is strongest at short antenna spacings and fades with increasing antenna spacing. The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) because the Galactic center is below the horizon.

@@

Diffuse emission outside the Galactic plane manifests in the primary field of view as a branch at $\tau>0$ and $\tau=0$ in the {\it off--zenith} and {\it zenith} pointings respectively. The former is seen at $\tau>0$ because the primary lobe of the power pattern is centered eastward of zenith, whereas in the latter it is centered at zenith. As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to the length of the baseline projected in the direction of the emission. Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at $\tau=0$, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0.75\arcdeg.

@@

The amplitude response of an interferometer to a point source is, to first order, flat cross baseline length. Since the primary lobe of the power pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with $\tau>0$ corresponding to the position of the primary lobe of the power pattern. In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at $\tau=0$ since the primary lobe of the power pattern is centered at zenith.

@@

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional {\bf delay power spectra} plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}. The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively. The horizon delay limits are shown as a pair of vertical dotted lines. The gray shaded area denotes the envelope of expected uncertainty in the {\bf delay power spectrum}. Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the % $\gtrsim$~70\% error owing to 
uncertainty in predicting the spectral index of compact foreground model, while thermal noise fluctuations dominate outside.

@@

The peak at $\tau=0$ (corresponding to the primary lobe in the power pattern) with a value of $\sim 10^7$--$10^8$~K$^2$(Mpc/$h$)$^3$, independent of antenna spacing, is predominantly determined by compact emission. The peak at $\tau=0$ from diffuse foreground model is $\sim 10^3$ times fainter and decreases rapidly with increase in antenna spacing. This is the response expected from different antenna spacings towards compact and diffuse emission.

@@

\begin{figure*}[htb]
\centering
% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figures/v1_0/sky_model_snapshot_0.eps}}
% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figures/v1_0/directional_delay_map_snapshot_0.eps}}
% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figures/v1_0/baseline_orientation_breakup_snapshot_0.eps}}
% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{f11a.eps}}
% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{f11b.eps}}
% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{f11c.eps}}
% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{fig12a.eps}}
% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{fig12b.eps}}
% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{fig12c.eps}}
\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figure9a.eps}}
\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figure9b.eps}}
\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figure9c.eps}}
\caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission (adopted from Figure~\ref{fig:sky-model-MWA}). The Galactic center is very prominent in diffuse emission on the west co--located with a bright secondary lobe of the power pattern. (b): Sky hemisphere mapped to delays observed on antenna spacings with northward ({\it top}) and eastward ({\it bottom}) orientations. Delays vary linearly with antenna spacing length. Color scale shown is for a 100~m antenna spacing. The bright Galactic center will appear at $\tau=0$ in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings. (c): Simulated delay spectra {\bf power} on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}). White lines denote horizon delay limits. The bright Galactic center is prominently visible close to negative horizon delay limit, especially on short eastward antenna spacings. These are also the most severely contaminated by foreground spillover. The northward antenna spacings, on the other hand, are the least contaminated.}
\label{fig:breakup}
\end{figure*}

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_before_after_nside_64_Tsys_95.0K.eps}
% \includegraphics[width=\linewidth]{f12.eps}
% \includegraphics[width=\linewidth]{fig13.eps}
\includegraphics[width=\linewidth]{figure10.eps}
\caption{Simulated delay spectra {\bf power} for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}). Discarded antenna spacings (black vertical stripes) have lengths $|\boldsymbol{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22.5\arcdeg. The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} is lowered by {\bf nearly two orders} of magnitude when short eastward antenna spacings are discarded. \label{fig:before-after}}
\end{figure}

@@

Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA {\bf to within expected uncertainties in foreground models}.

>>>>

Detection of 21~cm emission of H{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission. {\bf We investigate the signatures of wide--field measurements and an all--sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs. We demonstrate interferometric measurements are inherently sensitive to all scales, including the largest angular scales, owing to the nature of wide--field measurements. These wide--field effects are generic to all observations but antenna shapes impact their amplitudes substantially. A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array}. Comparing data from recent Murchison Widefield Array observations, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. We identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf that usually represent structures on small scales}. For signals entering through the primary field of view, compact emission dominates the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier delay space. Based on these results, we propose that selective down--weighting of data based on antenna spacing and time can mitigate foreground contamination substantially by a factor $\sim 100$ with negligible loss of sensitivity.

@@

{\bf In \S\ref{sec:delay-spectrum} we provide an overview of the delay spectrum technique. We investigate signatures generic to all wide--field measurements of EoR power spectra in \S\ref{sec:wide-field}. In \S\ref{sec:sim}, we present the foreground model and a variety of instrument models to rank antenna shapes based on foreground contamination. In \S\ref{sec:MWA}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are compared with the data and analyzed for foreground signatures. We report two important findings: foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique that minimizes foreground contamination using prior knowledge of the sky to down--weight adversely contaminated baselines}. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

We provide a short overview of the delay spectrum technique \citep{par12a,par12b}.

Interferometer array data known as {\it visibilities}, $V_b(f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis. If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_\textrm{i}(f)$ denotes instrumental bandpass weights, then $V_b(f)$ can be written as:
\begin{align}\label{eqn:obsvis}
  V_b(f) &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
\end{align}
where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector. This equation is valid {\bf in general, including wide--field measurements, and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}}.

@@

In order to express a quantity derived from $\tilde{V}_b(\tau)$ whose units are the same as that of the cosmological H{\sc i} power spectrum, we define the delay power spectrum:
\begin{align}\label{eqn:delay-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \\
  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
\end{align}
where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ is in units of K$^2$(Mpc/$h$)$^3$.

@@

{\bf In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s. Thus the MWA {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade}.

@@

In wide--field measurements, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view, a point also supported by Pober et al. (2015, in preparation). We use a foreground model that includes both diffuse and bright compact components.

@@

These criteria yield 100 objects from the SUMSS catalog and 250 objects from the NVSS catalog. Together with the diffuse foreground model, we obtain an all--sky foreground model consisting of both compact and diffuse emission. Figure~\ref{fig:sky-model-generic} shows the diffuse (top) and compact (bottom) foreground emission model used in our study. In this snapshot pointed towards zenith at 0.09~hours LST, the Galactic center in the diffuse model has just set in the West.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figure4.eps}
\includegraphics[width=\linewidth]{figure3.eps}
\caption{Foreground model at 185~MHz consisting of diffuse emission (top) in units of K and bright point sources (bottom) in units of Jy, visible during a snapshot at 0.09~hours LST. In the diffuse model, the Galactic center has just set in the West. Color scales are logarithmic. \label{fig:sky-model-generic}}
\end{figure}

@@

The power patterns of the aforementioned antenna geometries at 185~MHz for this zenith pointing are shown in Figure~\ref{fig:power-patterns}.

@@

We now use our simulations to analyze features in observed delay power spectrum obtained using the MWA instrument \citep{lon09,tin13}.

@@

MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013. The MWA is a 128--tile interferometer capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie. This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes. See \citet{bea12} for the tile layout. %{\bf At the observing frequency of 185~MHz, we use $A_\textrm{e}=10.5$~m$^2$}.

@@

{\bf The MWA is expected to be sensitive to the power spectrum of the H{\sc i} signal over the redshift range $6<z<10$ \citep{bow06,bea13,thy13}. Over 600~hours have been currently observed using the MWA, targeting science objectives outlined in \citet{bow13}.}

The MWA targets two primary low--foreground fields for reionization observations. Here, we focus on the field at RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$. The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of the dipoles in an MWA tile. The pointing system is optimized to points on a regular $\sim$7\arcdeg~ grid. The sky drifts across to the nearest available pointing, shifting between grid points (once every $\sim 30$~minutes). This process is repeated throughout the course of the observation $\approx 4.86$~hours.

The observations {\bf used} here were acquired on 2013~August~23. We have chosen two sections of duration 112 seconds each from this night for detailed study. These were chosen to provide a selection of possible foreground and instrumental conditions. As an example of a nominal observing setup we choose a zenith pointing; as an example of poor foreground conditions, we choose a pointing when the field is $\sim 2$~hours from zenith. This pointing has a significantly higher secondary lobe structure and is observed when the bright galactic center is well above the horizon.

@@

{\bf We estimate thermal noise, $\Delta V$, in the observed visibilities, $V_b(f)$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained from data after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
\begin{align}\label{eqn:Tsys}
  \Delta \tilde{V} &= \sqrt{N_\textrm{ch}}\,\Delta V \Delta f, \quad\textrm{and} \\
  \Delta V &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}}{A_\textrm{e}\sqrt{2\,\Delta f\,\Delta t}},
\end{align}
where, $\Delta f = 80$~kHz, $\Delta t = 112$~s, and $N_\textrm{ch}=\Delta B/\Delta f$ is the number of frequency channels. The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus yields a robust estimate of $T_\textrm{sys}$. We find the average system temperature to be $\sim 95$~K. Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.}

@@

Having shown that the simulation matches the data to the level of expected uncertainties, we proceed to examine in further detail the key signatures seen in simulated delay spectra. A number of factors are responsible for the characteristics noted in the delay spectra obtained from data and through simulations. In subsequent sections, we provide a detailed explanation of our results as a combination of these factors. Note that numerous features may overlap at different degrees of significance depending on combinations of parameters. We assign the features to their predominant causes. Secondly, we have used noiseless cases to clearly illustrate the observed foreground signatures. With the addition of noise in the visibilities, some of the weaker features may not be as prominently visible. Since the foreground signatures are far too numerous {\bf and subject to a multitude of parameters like} baseline length and orientation, power pattern, patch of sky under observation, and instrumental configuration, we highlight only the most notable features in the foreground delay power spectra.

@@

The most prominent signature seen in the {\it off--zenith} pointing (top left panel of Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power pattern. It appears as a bright branch near the negative delay horizon delay limit. This feature is strongest at short antenna spacings and fades with increasing antenna spacing. The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}) because the Galactic center has set below the horizon.

@@

Diffuse emission outside the Galactic plane manifests in the primary field of view as a branch at $\tau>0$ and $\tau=0$ in the {\it off--zenith} and {\it zenith} pointings respectively. The former is seen at $\tau>0$ because the primary lobe of the power pattern is centered eastward of zenith, whereas in the latter it is centered at zenith. As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to the length of the baseline projected in the direction of the emission. Thus, in the {\it zenith} pointing, the horizontal line at $\tau=0$, fades away on antenna spacings $|\boldsymbol{b}| \gtrsim 125$~m because the diffuse sky model is devoid of spatial structures on scales $\lesssim$~0.75\arcdeg.

@@

The amplitude response of an interferometer to a point source is, to first order, flat across baseline length. Since the primary field of view in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with $\tau>0$. In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at $\tau=0$ since the primary field of view is centered at zenith.

@@

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the {\bf delay power spectra} plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}. The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively. The horizon delay limits are shown as a pair of vertical dotted lines. The gray shaded area denotes the envelope of expected uncertainty in the {\bf delay power spectrum}. Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the % $\gtrsim$~70\% error owing to 
uncertainty in predicting the spectral index of compact foreground model, while thermal noise fluctuations dominate outside.

@@

The peak at $\tau=0$ (corresponding to the primary lobe in the power pattern) with a value of $\sim 10^7$--$10^8$~K$^2$(Mpc/$h$)$^3$, independent of antenna spacing, is predominantly determined by compact emission. The peak at $\tau=0$ from diffuse emission is $\sim 10^3$ times fainter and decreases rapidly with increase in antenna spacing. This is the response expected from different antenna spacings towards compact and diffuse emission.

@@

\begin{figure*}[htb]
\centering
% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figures/v1_0/sky_model_snapshot_0.eps}}
% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figures/v1_0/directional_delay_map_snapshot_0.eps}}
% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figures/v1_0/baseline_orientation_breakup_snapshot_0.eps}}
% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{f11a.eps}}
% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{f11b.eps}}
% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{f11c.eps}}
% \subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{fig12a.eps}}
% \subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{fig12b.eps}}
% \subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{fig12c.eps}}
\subfloat[][Sky model]{\label{fig:sky-model}\includegraphics[width=0.33\linewidth]{figure9a.eps}}
\subfloat[][Delay maps]{\label{fig:delay-map}\includegraphics[width=0.33\linewidth]{figure9b.eps}}
\subfloat[][Delay Spectra]{\label{fig:baseline-breakup}\includegraphics[width=0.33\linewidth]{figure9c.eps}}
\caption{(a): Sky model showing compact ({\it top}) and diffuse ({\it bottom}) emission (adopted from Figure~\ref{fig:sky-model-MWA}). The Galactic center is very prominent in diffuse emission on the west co--located with a bright secondary lobe of the power pattern. (b): Sky hemisphere mapped to delays observed on antenna spacings with northward ({\it top}) and eastward ({\it bottom}) orientations. Delays vary linearly with antenna spacing length. Color scale shown is for a 100~m antenna spacing. The bright Galactic center will appear at $\tau=0$ in northward antenna spacings and close to negative horizon delay limit on eastward antenna spacings. (c): Simulated delay {\bf power} spectra on antenna spacings oriented northward ({\it top}) and eastward ({\it bottom}). White lines denote horizon delay limits. The bright Galactic center is prominently visible close to negative horizon delay limit, especially on short eastward antenna spacings. These are also the most severely contaminated by foreground spillover. The northward antenna spacings, on the other hand, are the least contaminated.}
\label{fig:breakup}
\end{figure*}

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/baseline_screening_before_after_nside_64_Tsys_95.0K.eps}
% \includegraphics[width=\linewidth]{f12.eps}
% \includegraphics[width=\linewidth]{fig13.eps}
\includegraphics[width=\linewidth]{figure10.eps}
\caption{Simulated delay spectra {\bf power} for the {\it off--zenith} pointing with all antenna spacings included ({\it top}) and with short eastward antenna spacings discarded ({\it bottom}). Discarded antenna spacings (black vertical stripes) have lengths $|\boldsymbol{b}|<30$~m (leftward of vertical dashed line) and orientations $|\theta_\textrm{b}|<$~22.5\arcdeg. The spillover from the bright Galactic center near the negative horizon delay limit from the {\it foreground wedge} is lowered by a factor $\sim 100$ when short eastward antenna spacings are discarded. \label{fig:before-after}}
\end{figure}

@@

Simulating in {\bf many} important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA {\bf to within expected uncertainties in foreground models}.
